Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2023-07-11 chapters/Paper_two.tex
Minor changes

Wetting phenomena have an impact on areas as diverse as surface coatings, printing, plant treatment with pesticides, or even pandemics~\cite{RevModPhys.69.931, doi:10.1146/annurev-fluid-011212-140734, cassie1944wettability, Lenormand_1990,DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347,BergeronNature,BhardwajPoF2020}.

@@

leading to film rupture~\cite{RevModPhys.69.931, RevModPhys.81.1131} or wrinkling~\cite{DASILVASOBRINHO19991204}. 

@@

A successful mathematical modelling of the dynamics of thin liquid films is based on the thin-film equation~\cite{Reynolds, RevModPhys.69.931}

@@

the solid surface: for a no-slip condition~\cite{RevModPhys.69.931}, in particular, $M(h)=\frac{h^3}{3\mu}$, where $\mu$ is the fluid dynamic viscosity.

@@

\subsection{Fluid-substrate interactions: the contact angle}\label{subsec:fluid_substrate}
For the disjoining pressure $\Pi(h)$ we use the following 
expression\textcolor{black}{\cite{RevModPhys.81.1131, Peschka9275, PhysRevE.92.061002, RevModPhys.69.931}}:
\begin{equation}\label{eq:disjoin_p}
    \Pi(h) = \kappa(\theta) \left[\left(\frac{h_{\ast}}{h+h_c}\right)^n - \left(\frac{h_{\ast}}{h+h_c}\right)^m\right],
\end{equation}
where the prefactor $\kappa(\theta) = \gamma(1-\cos(\theta))\frac{(n-1)(m-1)}{(n-m)h_{\ast}}$ is related to the Hamaker constant by
$\mathcal{A} = 6 \pi h_{\ast}^3 \kappa(\theta)$.
$\theta$ is the contact angle, which effectively encodes the interfacial energies of the three-phase system and represents, therefore, a measure of the hydrophilicity/hydrophobicity of the substrate.
\textcolor{black}{Throughout the paper we restrict ourselves to low contact angles, $\theta < 1$,
so to comply with the lubrication approximation, which requires the derivative of the height field to be small, $\partial_x h(x) \ll 1$, and hence limits
the values of $\theta$ (since, at the contact point, $\partial_x h \sim \tan(\theta) \sim \theta$).}
\textcolor{black}{To ensure numerical stability we introduce a regularizing term $h_c$; the value $h_{\ast} - h_c > 0$, 
at which the disjoining pressure vanishes, sets the thickness of the precursor film covering the substrate.}

>>>>

Wetting phenomena have an impact on areas as diverse as surface coatings, printing, plant treatment with pesticides, or even pandemics~\cite{oronLongscaleEvolutionThin1997, doi:10.1146/annurev-fluid-011212-140734, cassieWettabilityPorousSurfaces1944, Lenormand_1990,DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347,BergeronNature,BhardwajPoF2020}.

@@

leading to film rupture~\cite{oronLongscaleEvolutionThin1997, RevModPhys.81.1131} or wrinkling~\cite{dasilvasobrinhoStudyDefectsUltrathin1999}. 

@@

A successful mathematical modelling of the dynamics of thin liquid films is based on the thin-film equation~\cite{Reynolds, oronLongscaleEvolutionThin1997}

@@

the solid surface: for a no-slip condition~\cite{oronLongscaleEvolutionThin1997}, in particular, $M(h)=\frac{h^3}{3\mu}$, where $\mu$ is the fluid dynamic viscosity.

@@

\subsection{Fluid-substrate interactions: the contact angle}\label{subsec:fluid_substrate}
For the disjoining pressure $\Pi(h)$ we use the following 
expression\textcolor{black}{\cite{RevModPhys.81.1131, Peschka9275, PhysRevE.92.061002, oronLongscaleEvolutionThin1997}}:
\begin{equation}\label{eq:disjoin_p}
    \Pi(h) = \kappa(\theta) \left[\left(\frac{h_{\ast}}{h+h_c}\right)^n - \left(\frac{h_{\ast}}{h+h_c}\right)^m\right],
\end{equation}
where the prefactor $\kappa(\theta) = \gamma(1-\cos(\theta))\frac{(n-1)(m-1)}{(n-m)h_{\ast}}$ is related to the Hamaker constant by
$\mathcal{A} = 6 \pi h_{\ast}^3 \kappa(\theta)$.
$\theta$ is the contact angle, which effectively encodes the interfacial energies of the three-phase system and represents, therefore, a measure of the hydrophilicity/hydrophobicity of the substrate.
\textcolor{black}{Throughout the paper we restrict ourselves to low contact angles, $\theta < 1$,
so to comply with the lubrication approximation, which requires the derivative of the height field to be small, $\partial_x h(x) \ll 1$, and hence limits
the values of $\theta$ (since, at the contact point, $\partial_x h \sim \tan(\theta) \sim \theta$).}
\textcolor{black}{To ensure numerical stability we introduce a regularizing term $h_c$; the value $h_{\ast} - h_c > 0$, 
at which the disjoining pressure vanishes, sets the thickness of the precursor film covering the substrate.}

### 2023-07-11 chapters/Theory.tex
Minor changes

For the momentum equation the same line of argument is applied, however momentum can change for different reasons.
First being that momentum is advected from or into the volume $V_0$.
Pressure on the other hand can be a source of momentum and of course forces that are applied to the fluid are as well sources~\cite{krueger2017}.
Collecting these terms and writing down an equation for the momentum one gets,
\begin{equation}\label{eq:navier_stokes_1}
    \partial_t \int_{V_0}\rho\mathbf{u}\diff V = -\oint_{\partial V_0}(\rho\mathbf{u}\mathbf{u})\cdot\diff\textbf{A} - \oint_{\partial V_0} p\diff\textbf{A} + \int_{V_0} \textbf{F}\diff V, 
\end{equation}
where $\mathbf{u}\mathbf{u}$ is used to denote the outer product.
Using the divergence theorem again to reformulate everything in terms of volume integrals and dropping the integration the above equation becomes
\begin{equation}\label{eq:navier_stokes_2}
    \partial_t(\rho\mathbf{u}) = -\nabla\cdot(\rho\mathbf{u}\mathbf{u}) - \nabla p + \textbf{F}.
\end{equation}
This however is not yet the Navier-Stokes equation, in fact this is the Euler equation~\cite{batchelor2000introduction}.
In the derivation there is one term missing, because the momentum of a fluid can also change due to internal friction, or put simply, due to viscosity.
This viscous contributions are collected in the viscous stress tensor $\hat{\sigma}$ as
\begin{equation}\label{eq:stress_tens}
    \hat{\sigma}_{ij} = \mu\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i}\right) + \lambda\delta_{ij}\frac{\partial u_k}{\partial x_k},
\end{equation}
Actually this tensor introduces a lot of the complexity of the Navier-Stokes equation as in principle neither $\mu$ nor $\lambda$ need to be constants and in fact they are tensors themselves.

@@

\subsection{Scaling analysis}
\label{subssec:scaling_shallow water}
Starting with the system specific approximations, in Eq.~(\ref{eq:thickness_length}) a so-called scaling parameter $\varepsilon$ has been introduced. 
Using this parameter to rescale the velocities, because we know that the horizontal velocities are large as compared to the vertical
\begin{equation}\label{eq:vel_scaling}
    u_x = \tilde{u}_x,\quad u_y = \tilde{u}_y,\quad u_z = \varepsilon\tilde{u}_z.
\end{equation}
For an incompressible fluid with vanishing density gradient Eq.~(\ref{eq:cont_3}) therefore becomes
\begin{equation}\label{eq:cont_nondim}
    \nabla\cdot\mathbf{u} = \partial_x u_x + \partial_y u_y + \varepsilon\partial_z u_z = 0.
\end{equation}
To satisfy that all terms of this equation are the same order in $\varepsilon$ the coordinates need to be rescaled as well. 
This choice however is not unique and offers two possibilities~\cite{james2019towards}.
There is on the one hand the long wave approximation~\cite{RevModPhys.69.931},
\begin{equation}\label{eq:long_wave_scaling}
    x = \frac{\tilde{x}}{\varepsilon},\quad y = \frac{\tilde{y}}{\varepsilon},\quad z = \tilde{z},\quad t = \frac{\tilde{t}}{\varepsilon},
\end{equation}
which again is encountered in Sec.~\ref{subsec:thin_film_scaling} and tries to smooth out spatial and temporal variations.
On the other hand is the thin layer scaling
\begin{equation}\label{eq:scaling_thin_layer}
    x = \tilde{x},\quad y = \tilde{y},\quad z = \varepsilon\tilde{z},\quad t = \tilde{t},
\end{equation}
which correspond to the classical scaling of boundary layer theory~\cite{milne1996theoretical}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/Simple_flow.pdf}
    \caption{a) Schematic, cut through the center, velocity profile of a pipe flow with wall friction and no slip boundary condition.
    The flow velocity $u_x$ can be calculated using Eq.~(\ref{eq:pois_vel_profile}).
    b) Similar case in terms of a shallow water geometry. 
    The upper wall is replaced with a pressure free boundary while the lower wall induces friction according to a no-slip boundary condition.
    }
    \label{fig:pois_vel_profile}
\end{figure}
\subsection{Poiseuille flow}
The Poiseuille flow is one amongst those few problems that can be solved analytically~\cite{sutera1993history}.
To have an idea about the flow, one can think about a circular pipe.
The pipe is fully filled with a (Newtonian) liquid, see Fig.~\ref{fig:pois_vel_profile} for an oversimplified illustration.
By an external pressure gradient, for example a pump, a continuous flow is generated.
The velocity profile of this flow can be explained with the equation~\cite{batchelor2000introduction, krueger2017}
\begin{equation}\label{eq:pois_vel_profile}
    u_x(y) = -\frac{\partial_x p}{2\mu}y(y-d),
\end{equation}
where $d$ the diameter of the pipe is.
The profile is parabolic with vanishing $u_x$ at the walls and a maximum at $d/2$. 
Thus the above equation defines the balance between driving force, here the pressure difference, and the friction at the walls.

@@

\section{Lubrication Theory and Thin Films}
\label{sec:thin_films}
Thin liquid films are an interesting problem to study.
Their dynamics is seemingly easy to explain but bare a lot of complexity.
Seemingly easy because everyone has an understanding of soap bubbles and their rather complex bursting mechanics or has seen a rain drop on a plant leaf as motivated in the introduction, Chap.~\ref{chapter:intro}.
Complexity is generated due to the shape of the equation that describes the thin film dynamics, usually a non linear forth order partial differential equation of the film thickness.
When the thickness of a film becomes small, e.g. only a few tens of nanometers the complexity can even be enhanced. 
Complex behaviour can arise due to thermal fluctuations or strictly speaking the mathematical unsolvable problem of moving contact lines e.g. when a wet to dry transition happens.
Similar to the above section's scaling argument and the long wave theory, the Navier-Stokes equation can be reduced to the thin film equation. 
Although with the difference to the shallow water system, see Eqs.~(\ref{eq:shallow_water}), that the thin film equation does only admit a continuity like equation and no accompanying momentum equation~\cite{RevModPhys.69.931, RevModPhys.57.827, RevModPhys.81.1131}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/contact_line.pdf}
    \caption{Liquid droplet placed on a smooth substrate.
    The three phase contact line which is the interface between the vapor, the solid and liquid phase is displayed as thick black line.}
    \label{fig:contact_line_drop}
\end{figure}
Similar to the shallow water system, the thin film equation only considers the dynamics of a thin liquid \textit{layer} placed on top of a yet undefined surface or substrate.
The defining equation that describes the dynamics of a thin liquid film can be written as~\cite{THIELE2014399, RevModPhys.69.931, RevModPhys.81.1131, RevModPhys.81.739} 
\begin{equation}\label{eq:simple_thin_film}
    \partial_t h = \vec{\nabla}\cdot\left(M(h)\vec{\nabla} p\right).
\end{equation}
Stating that the height can only change if there is a pressure gradient which acts on the film.
In Eq.~(\ref{eq:simple_thin_film}) two unknowns are introduced.
First being the mobility $M(h)$ which determines the exact velocity boundary condition for $h=0$.
Using the no-slip boundary condition, see Eq.~(\ref{eq:no-slip_sw1}), we have 
\begin{equation}\label{eq:no-slip-mobility}
    M(h) = \frac{h^3}{3\mu},
\end{equation}
with $\mu$ being the (dynamic) viscosity, therefore the mobility is inversely proportional to the viscosity.
Of course the velocity boundary condition can be relaxed which is often desired e.g. for contact line movement.
For problems such as dewetting\footnote{Dewetting usually refers to an unstable film which by so-called film rupture recedes into separated droplets. Depending on the film's properties there different modes of dewetting can be identified, such as spinodal or hole nucleation.} it is actually a known approximation to relax the velocity boundary condition at the liquid solid interface~\cite{M_nch_2005, munch2005lubrication, PhysRevLett.95.127801}.
The term contact line or to be precise three-phase contact line defines the single one dimensional interface that connects all three phases involved in the problem, as displayed in Fig.~\ref{fig:contact_line_drop}.
The addition of the parameter $\delta$ makes it possible to interpolate the height of vanishing fluid velocity into the substrate.
Using this interpolation the mobility function Eq.~(\ref{eq:no-slip-mobility}) can be extended towards a slip like behaviour in the proximity of $h\approx 0$ as
\begin{equation}\label{eq:slip_mobility}
    M(h) = \mu^{-1}\left(\frac{h^3}{3} + \delta h^2\right).
\end{equation}

@@

Inserting the rescaled quantities the continuity equation becomes
\begin{equation}\label{eq:cont_thin_film_1}
     \nabla\cdot\mathbf{u} = \partial_{\tilde{x}} \tilde{u}_x + \partial_{\tilde{y}} \tilde{u}_y + \frac{V L}{U H}\partial_{\tilde{z}} \tilde{u}_z = 0,
\end{equation}
which means that $V = U H/L$ to ensure all terms are of the same order in $\varepsilon$.
Upon inserting them into the Navier-Stokes equation we get for the x-component only
\begin{align}
     \rho L \partial_{\tilde{t}}\tilde{u}_x &+ \rho\left[\frac{U^2}{L}\left(\tilde{u}_x\partial_{\tilde{x}}\tilde{u}_x + \tilde{u}_y\partial_{\tilde{y}}\tilde{u}_x\right) + \frac{U V}{H}\tilde{u}_z\partial_{\tilde{z}}\tilde{u}_x\right] = \nonumber \\ &-\frac{\mu U}{H^2} \partial_{\tilde{x}} \tilde{p} + \mu\left[\frac{U}{L^2}(\partial^2_{\tilde{x}}\tilde{u}_x + \partial^2_{\tilde{y}}\tilde{u}_x) + \frac{U}{H^2}\partial^2_{\tilde{z}}\tilde{u}_x\right].
\end{align}
For the $y$ and $z$ component we refer to the review by Oron et al.~\cite{RevModPhys.69.931}.
Although the equation is now non-dimensional it is instructive to multiply both sides with $H^2/\mu U$ and collect the characteristic quantities in dimensional less numbers, here the Reynolds number, see Eq.~(\ref{eq:Re_and_Fr}), with $L = H, v_0 = U$ and $\varepsilon$  
\begin{equation}
    \varepsilon Re \left(\partial_{\tilde{t}}\tilde{u}_x + \tilde{u}_x\partial_{\tilde{x}}\tilde{u}_x + \tilde{u}_y\partial_{\tilde{y}}\tilde{u}_x + \tilde{u}_z\partial_{\tilde{z}}\tilde{u}_x \right) = -\partial_{\tilde{x}}\tilde{p} + \varepsilon^2\left(\partial^2_{\tilde{x}}\tilde{u}_x + \partial^2_{\tilde{y}}\tilde{u}_x\right) + \partial^2_{\tilde{z}}\tilde{u}_x. 
\end{equation}
Both the Reynolds number as well as $\varepsilon$ are by definition small. 
Therefore the terms with higher order in $\varepsilon$ are at most subleading.
Collecting only the term of $O(1)$ and performing the same scaling for the two other directions leads to the non-dimensional momentum equation of a thin film on a solid substrate
\begin{align}
    -\partial_{\tilde{x}}\tilde{p} + \partial^2_{\tilde{z}}\tilde{u}_x &= 0,  \label{eq:thin_film_pressure_x}\\
    -\partial_{\tilde{y}}\tilde{p} + \partial^2_{\tilde{z}}\tilde{u}_y &= 0,  \label{eq:thin_film_pressure_y}\\
    \partial_{\tilde{z}}\tilde{p} &= 0. \label{eq:thin_film_pressure_z}
\end{align}
where for the last of the above equations the analogy to the shallow water model can be drawn using Eq.~(\ref{eq:hydrostat_sw_euler}) which states that the pressure gradient does not depend on $z$.

>>>>

For the momentum equation the same line of argument is applied, however momentum can change for different reasons.
First being that momentum is advected from or into the volume $V_0$.
Pressure on the other hand can be a source of momentum and of course forces that are applied to the fluid are as well sources~\cite{krugerLatticeBoltzmannMethod2017}.
Collecting these terms and writing down an equation for the momentum one gets,
\begin{equation}\label{eq:navier_stokes_1}
    \partial_t \int_{V_0}\rho\mathbf{u}\diff V = -\oint_{\partial V_0}(\rho\mathbf{u}\mathbf{u})\cdot\diff\textbf{A} - \oint_{\partial V_0} p\diff\textbf{A} + \int_{V_0} \textbf{F}\diff V, 
\end{equation}
where $\mathbf{u}\mathbf{u}$ is used to denote the outer product.
Using the divergence theorem again to reformulate everything in terms of volume integrals and dropping the integration the above equation becomes
\begin{equation}\label{eq:navier_stokes_2}
    \partial_t(\rho\mathbf{u}) = -\nabla\cdot(\rho\mathbf{u}\mathbf{u}) - \nabla p + \textbf{F}.
\end{equation}
This however is not yet the Navier-Stokes equation, in fact this is the Euler equation~\cite{batchelor2000introduction}.
In the derivation there is one term missing, because the momentum of a fluid can also change due to internal friction, or put simply, due to viscosity.
This viscous contributions are collected in the viscous stress tensor $\hat{\sigma}$ as
\begin{equation}\label{eq:stress_tens}
    \hat{\sigma}_{ij} = \mu\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i}\right) + \lambda\delta_{ij}\frac{\partial u_k}{\partial x_k},
\end{equation}
Actually this tensor introduces a lot of the complexity of the Navier-Stokes equation as in principle neither $\mu$ nor $\lambda$ need to be constants and in fact they are tensors themselves.

@@

\subsection{Scaling analysis}
\label{subssec:scaling_shallow water}
Starting with the system specific approximations, in Eq.~(\ref{eq:thickness_length}) a so-called scaling parameter $\varepsilon$ has been introduced. 
Using this parameter to rescale the velocities, because we know that the horizontal velocities are large as compared to the vertical
\begin{equation}\label{eq:vel_scaling}
    u_x = \tilde{u}_x,\quad u_y = \tilde{u}_y,\quad u_z = \varepsilon\tilde{u}_z.
\end{equation}
For an incompressible fluid with vanishing density gradient Eq.~(\ref{eq:cont_3}) therefore becomes
\begin{equation}\label{eq:cont_nondim}
    \nabla\cdot\mathbf{u} = \partial_x u_x + \partial_y u_y + \varepsilon\partial_z u_z = 0.
\end{equation}
To satisfy that all terms of this equation are the same order in $\varepsilon$ the coordinates need to be rescaled as well. 
This choice however is not unique and offers two possibilities~\cite{james2019towards}.
There is on the one hand the long wave approximation~\cite{oronLongscaleEvolutionThin1997},
\begin{equation}\label{eq:long_wave_scaling}
    x = \frac{\tilde{x}}{\varepsilon},\quad y = \frac{\tilde{y}}{\varepsilon},\quad z = \tilde{z},\quad t = \frac{\tilde{t}}{\varepsilon},
\end{equation}
which again is encountered in Sec.~\ref{subsec:thin_film_scaling} and tries to smooth out spatial and temporal variations.
On the other hand is the thin layer scaling
\begin{equation}\label{eq:scaling_thin_layer}
    x = \tilde{x},\quad y = \tilde{y},\quad z = \varepsilon\tilde{z},\quad t = \tilde{t},
\end{equation}
which correspond to the classical scaling of boundary layer theory~\cite{milne1996theoretical}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/Simple_flow.pdf}
    \caption{a) Schematic, cut through the center, velocity profile of a pipe flow with wall friction and no slip boundary condition.
    The flow velocity $u_x$ can be calculated using Eq.~(\ref{eq:pois_vel_profile}).
    b) Similar case in terms of a shallow water geometry. 
    The upper wall is replaced with a pressure free boundary while the lower wall induces friction according to a no-slip boundary condition.
    }
    \label{fig:pois_vel_profile}
\end{figure}
\subsection{Poiseuille flow}
The Poiseuille flow is one amongst those few problems that can be solved analytically~\cite{sutera1993history}.
To have an idea about the flow, one can think about a circular pipe.
The pipe is fully filled with a (Newtonian) liquid, see Fig.~\ref{fig:pois_vel_profile} for an oversimplified illustration.
By an external pressure gradient, for example a pump, a continuous flow is generated.
The velocity profile of this flow can be explained with the equation~\cite{batchelor2000introduction, krugerLatticeBoltzmannMethod2017}
\begin{equation}\label{eq:pois_vel_profile}
    u_x(y) = -\frac{\partial_x p}{2\mu}y(y-d),
\end{equation}
where $d$ the diameter of the pipe is.
The profile is parabolic with vanishing $u_x$ at the walls and a maximum at $d/2$. 
Thus the above equation defines the balance between driving force, here the pressure difference, and the friction at the walls.

@@

\section{Lubrication Theory and Thin Films}
\label{sec:thin_films}
Thin liquid films are an interesting problem to study.
Their dynamics is seemingly easy to explain but bare a lot of complexity.
Seemingly easy because everyone has an understanding of soap bubbles and their rather complex bursting mechanics or has seen a rain drop on a plant leaf as motivated in the introduction, Chap.~\ref{chapter:intro}.
Complexity is generated due to the shape of the equation that describes the thin film dynamics, usually a non linear forth order partial differential equation of the film thickness.
When the thickness of a film becomes small, e.g. only a few tens of nanometers the complexity can even be enhanced. 
Complex behaviour can arise due to thermal fluctuations or strictly speaking the mathematical unsolvable problem of moving contact lines e.g. when a wet to dry transition happens.
Similar to the above section's scaling argument and the long wave theory, the Navier-Stokes equation can be reduced to the thin film equation. 
Although with the difference to the shallow water system, see Eqs.~(\ref{eq:shallow_water}), that the thin film equation does only admit a continuity like equation and no accompanying momentum equation~\cite{oronLongscaleEvolutionThin1997, RevModPhys.57.827, RevModPhys.81.1131}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/contact_line.pdf}
    \caption{Liquid droplet placed on a smooth substrate.
    The three phase contact line which is the interface between the vapor, the solid and liquid phase is displayed as thick black line.}
    \label{fig:contact_line_drop}
\end{figure}
Similar to the shallow water system, the thin film equation only considers the dynamics of a thin liquid \textit{layer} placed on top of a yet undefined surface or substrate.
The defining equation that describes the dynamics of a thin liquid film can be written as~\cite{thielePatternedDepositionMoving2014, oronLongscaleEvolutionThin1997, RevModPhys.81.1131, RevModPhys.81.739} 
\begin{equation}\label{eq:simple_thin_film}
    \partial_t h = \vec{\nabla}\cdot\left(M(h)\vec{\nabla} p\right).
\end{equation}
Stating that the height can only change if there is a pressure gradient which acts on the film.
In Eq.~(\ref{eq:simple_thin_film}) two unknowns are introduced.
First being the mobility $M(h)$ which determines the exact velocity boundary condition for $h=0$.
Using the no-slip boundary condition, see Eq.~(\ref{eq:no-slip_sw1}), we have 
\begin{equation}\label{eq:no-slip-mobility}
    M(h) = \frac{h^3}{3\mu},
\end{equation}
with $\mu$ being the (dynamic) viscosity, therefore the mobility is inversely proportional to the viscosity.
Of course the velocity boundary condition can be relaxed which is often desired e.g. for contact line movement.
For problems such as dewetting\footnote{Dewetting usually refers to an unstable film which by so-called film rupture recedes into separated droplets. Depending on the film's properties there different modes of dewetting can be identified, such as spinodal or hole nucleation.} it is actually a known approximation to relax the velocity boundary condition at the liquid solid interface~\cite{M_nch_2005, munch2005lubrication, PhysRevLett.95.127801}.
The term contact line or to be precise three-phase contact line defines the single one dimensional interface that connects all three phases involved in the problem, as displayed in Fig.~\ref{fig:contact_line_drop}.
The addition of the parameter $\delta$ makes it possible to interpolate the height of vanishing fluid velocity into the substrate.
Using this interpolation the mobility function Eq.~(\ref{eq:no-slip-mobility}) can be extended towards a slip like behaviour in the proximity of $h\approx 0$ as
\begin{equation}\label{eq:slip_mobility}
    M(h) = \mu^{-1}\left(\frac{h^3}{3} + \delta h^2\right).
\end{equation}

@@

Inserting the rescaled quantities the continuity equation becomes
\begin{equation}\label{eq:cont_thin_film_1}
     \nabla\cdot\mathbf{u} = \partial_{\tilde{x}} \tilde{u}_x + \partial_{\tilde{y}} \tilde{u}_y + \frac{V L}{U H}\partial_{\tilde{z}} \tilde{u}_z = 0,
\end{equation}
which means that $V = U H/L$ to ensure all terms are of the same order in $\varepsilon$.
Upon inserting them into the Navier-Stokes equation we get for the x-component only
\begin{align}
     \rho L \partial_{\tilde{t}}\tilde{u}_x &+ \rho\left[\frac{U^2}{L}\left(\tilde{u}_x\partial_{\tilde{x}}\tilde{u}_x + \tilde{u}_y\partial_{\tilde{y}}\tilde{u}_x\right) + \frac{U V}{H}\tilde{u}_z\partial_{\tilde{z}}\tilde{u}_x\right] = \nonumber \\ &-\frac{\mu U}{H^2} \partial_{\tilde{x}} \tilde{p} + \mu\left[\frac{U}{L^2}(\partial^2_{\tilde{x}}\tilde{u}_x + \partial^2_{\tilde{y}}\tilde{u}_x) + \frac{U}{H^2}\partial^2_{\tilde{z}}\tilde{u}_x\right].
\end{align}
For the $y$ and $z$ component we refer to the review by Oron et al.~\cite{oronLongscaleEvolutionThin1997}.
Although the equation is now non-dimensional it is instructive to multiply both sides with $H^2/\mu U$ and collect the characteristic quantities in dimensional less numbers, here the Reynolds number, see Eq.~(\ref{eq:Re_and_Fr}), with $L = H, v_0 = U$ and $\varepsilon$  
\begin{equation}
    \varepsilon Re \left(\partial_{\tilde{t}}\tilde{u}_x + \tilde{u}_x\partial_{\tilde{x}}\tilde{u}_x + \tilde{u}_y\partial_{\tilde{y}}\tilde{u}_x + \tilde{u}_z\partial_{\tilde{z}}\tilde{u}_x \right) = -\partial_{\tilde{x}}\tilde{p} + \varepsilon^2\left(\partial^2_{\tilde{x}}\tilde{u}_x + \partial^2_{\tilde{y}}\tilde{u}_x\right) + \partial^2_{\tilde{z}}\tilde{u}_x. 
\end{equation}
Both the Reynolds number as well as $\varepsilon$ are by definition small. 
Therefore the terms with higher order in $\varepsilon$ are at most subleading.
Collecting only the term of $O(1)$ and performing the same scaling for the two other directions leads to the non-dimensional momentum equation of a thin film on a solid substrate
\begin{align}
    -\partial_{\tilde{x}}\tilde{p} + \partial^2_{\tilde{z}}\tilde{u}_x &= 0,  \label{eq:thin_film_pressure_x}\\
    -\partial_{\tilde{y}}\tilde{p} + \partial^2_{\tilde{z}}\tilde{u}_y &= 0,  \label{eq:thin_film_pressure_y}\\
    \partial_{\tilde{z}}\tilde{p} &= 0. \label{eq:thin_film_pressure_z}
\end{align}
where for the last of the above equations the analogy to the shallow water model can be drawn using Eq.~(\ref{eq:hydrostat_sw_euler}) which states that the pressure gradient does not depend on $z$.

### 2023-07-11 main.tex
More structure a little more text
Iteration with Jens comments
Updates from Overleaf
jens_comments_work
Added a readme
gh action and prolog
Minor changes

\documentclass[titlepage,a4paper,twoside,12pt,bibliography=totoc,BCOR=6.5mm]{scrreprt}

\usepackage{main}
% use biber for build
\addbibresource{bibliography/references.bib}%

\author{Stefan Zitz}%
\title{Dynamics of thin liquid film flows studied with a lattice Boltzmann method}%
\date{\today}%

@@

\begin{document}%
%
\pagenumbering{alph}%
\pagestyle{empty}%
\include{others/titlepage}%
%
\cleardoublepage%
%
\cleardoublepage%
%
\include{others/abstract}%
%
\cleardoublepage%
%
\pagestyle{scrheadings}%
\pagenumbering{roman}%
\setcounter{page}{1}%
%
\tableofcontents%
%
\cleardoublepage%
%
\listoffigures%
%
\cleardoublepage%
%
\pagenumbering{arabic}%
\setcounter{page}{1}%
%
%
\include{chapters/Introduction}%
\include{chapters/Theory}
\include{chapters/Method}
%
%
\cleardoublepage%
\nocite{*}%
\printbibliography%
%
%
\cleardoublepage%
%
\appendix%
\renewcommand{\thepage}{\Alph{chapter}\ \arabic{page}}%
%
\setcounter{page}{1}%
\include{appendix/example}%
%
\end{document}

>>>>

\documentclass[titlepage,a4paper,twoside,12pt,headheight=30pt,bibliography=totoc,BCOR=6.5mm]{scrreprt}

\usepackage{main}
% use biber for build
\addbibresource{bibliography/references.bib}%
\addbibresource{bibliography/Thesis_Swalbe.bib}%

\author{Stefan Zitz}%
\title{Dynamics of thin liquid film flows studied with the lattice Boltzmann method}%
\date{\today}%

@@

\begin{document}%
%
\pagenumbering{alph}%
\pagestyle{empty}%
\include{others/titlepage}%
%
\cleardoublepage%
%
\cleardoublepage%
%
\include{others/abstract}%
%
\cleardoublepage%
%
\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}

\pagestyle{scrheadings}%
\pagenumbering{roman}%
\setcounter{page}{1}%
\include{others/prolog}%%
\tableofcontents%
%
\cleardoublepage%
%
\listoffigures%
%
\cleardoublepage%
%

\pagenumbering{arabic}%
\setcounter{page}{1}%
%
%
\include{chapters/Introduction}%
\include{chapters/Theory}
\include{chapters/Method}
\include{chapters/Paper_four}
\include{chapters/Paper_one}
\include{chapters/Paper_two}
% \include{chapters/Paper_five}
\include{chapters/Paper_three}
\include{chapters/Conclusion}
%
%
%
\cleardoublepage%
\nocite{*}%
\printbibliography%
%
%
\cleardoublepage%

\appendix%
\renewcommand{\thepage}{\Alph{chapter}\ \arabic{page}}%
%
\cleardoublepage%
%
\include{others/acknowledgements}
%
\setcounter{page}{1}%
%\include{appendix/example}%
%
\end{document}
